\name{}: machine learning over interaction networks and ontology-based background knowledge predicts drug--target interactions

### 2021-07-04 paper/main.tex
.

To the best of our knowledge, \name{} is the first DTI prediction
model that propagates ontology-based features of protein interaction
networks; while a similar approach of combining ontology embeddings
with interaction networks has previously been used for analyzing gene
expression \citet{Trebacz2020}, \name{} extends this method to DTI
prediction. Further, \name{} is novel in that it explicitly integrates
``bottom-up'' and ``top-down'' features using a graph representation
of interactions between proteins; while there are other DTI prediction
methods that combine these features \cite{Gottlieb2011}, \name{}
exploits the ability to integrate heterogeneous features using graphs
and the ability to utilize this graph in machine learning through the
use of graph neural networks.

@@

Several of the biases we identify in evaluating DTI prediction methods
have been observed previously. The performance difference based on how
training and evaluation data is split (by interaction pair, by drug,
or by protein) has been demonstrated before using a
\todo{MacroAUC/MicroAUC???} measure \cite{}; we further extend on
these results by introducing performance measures based on micro
averages (Micro AUC$_p$ and Micro AUC$_d$) to further illustrate how
prediction performance changes when evaluation data is imbalanced. We
have further extended on prior results by introducing a ``na\"ive''
classifier that explicitly exploits one data bias to make predictions,
illustrating that this bias has a significant impact on DTI
methods. The bias we identify with the ``na\"ive'' classifier is
similar to a previous bias found in gene networks when using methods
that rely on the guilt-by-association principle
\cite{10.1371/journal.pcbi.1002444} but which has, to our knowledge,
not been demonstrated in the context of DTI prediction.

>>>>

To the best of our knowledge, \name{} is the first DTI prediction
model that propagates ontology-based features of protein interaction
networks; while a similar approach of combining ontology embeddings
with interaction networks has previously been used for analyzing gene
expression \cite{Trebacz2020}, \name{} extends this method to DTI
prediction. Further, \name{} is novel in that it explicitly integrates
``bottom-up'' and ``top-down'' features using a graph representation
of interactions between proteins; while there are other DTI prediction
methods that combine these features \cite{Gottlieb2011}, \name{}
exploits the ability to integrate heterogeneous features using graphs
and the ability to utilize this graph in machine learning through the
use of graph neural networks.

@@

Several of the biases we identify in evaluating DTI prediction methods
have been observed previously. The performance difference based on how
training and evaluation data is split (by interaction pair, by drug,
or by protein) has been demonstrated before using a
\todo{MacroAUC/MicroAUC???} measure \cite{}; we further extend on
these results by introducing performance measures based on micro
averages (Micro AUC$_p$ and Micro AUC$_d$) to further illustrate how
prediction performance changes when evaluation data is imbalanced. We
have further extended on prior results by introducing a ``na\"ive''
classifier that explicitly exploits one data bias to make predictions,
illustrating that this bias has a significant impact on DTI
methods. The bias we identify with the ``na\"ive'' classifier is
similar to a previous bias found in gene networks when using methods
that rely on the guilt-by-association principle
\cite{Gillis2012} but which has, to our knowledge,
not been demonstrated in the context of DTI prediction.
